---
title: "Nathalie"
format: html
---

## The Study
This analysis aimed to reproduce the results from Lépissier, Alice & Mildenberger, Matto. (2021). Unilateral climate policies can substantially reduce national carbon pollution. Climatic Change. 166. 10.1007/s10584-021-03111-2. The authors employed a synthetic control method (SCM) to estimate the causal effect of the UK’s 2001 Climate Change Programme (CCP) on the country’s carbon emissions. Their findings demonstrate that CCP reduced the UK's carbon emissions. CO2 emissions per capita were 9.8% lower in 2005 relative to what they would have been if the CCP had not been passed, demonstrating that unilateral climate policies can still reduce carbon emissions even in the absence of a binding global climate agreement.

The 2001 UK CCP is a complex reform that included a carbon tax on large-scale energy users, industry-negotiated exemptions from the tax for meeting reduction targets, and a voluntary emissions trading scheme. It was one of the first comprehensive climate reform packages passed globally, in advance of action by most other Organisation for Economic Co-operation and Development (OECD) countries. Importantly, there is a limitation to evaluating the post-policy enactment effect, as the EU emissions trading system (EU ETS) was enacted in 2005.

This positions the UK's CCP to be evaluated against a wide pool of donor countries by pre-treatment carbon emissions. The study compares 

## Data

-   data: all countries

-   data_OECD: similar economic conditions

-   data_OECD_HIC: + high income countries

-   data_OECD_HIC_UMC: + upper middle countries

# Load libraries
```{r}
#| output: false
library(tidysynth) # for synthetic control
library(tidyverse)
library(patchwork) # for combining plots
library(here)
```

## Source function to import .Rdata files from environment
```{r}
source(here("R", "load_rdata_files.R"))

load_rdata_files()
```

## Declare the Synthetic Control - FIRST TRY (#1)
```{r}
# data_OECD
scm_data_OECD <- data_OECD %>%
  synthetic_control(
    outcome = `EN.ATM.CO2E.PC`, 
    unit = country,
    time = year,
    i_unit = "United Kingdom",
    i_time = 2001,
    generate_placebos = TRUE) %>%
  
  generate_predictor(  # Predictors (pre-treatment averages)
    time_window = 1980:2001,
    mean.preT = mean(mean.preT, na.rm = TRUE)) %>%
  
  generate_weights(
    optimization_window = 1980:2001,
    margin_ipop = 0.02,
    sigf_ipop = 7,
    bound_ipop = 6) %>%
  generate_control() # Generate synthetic control


# data_OECD_HIC
scm_data_OECD_HIC <- data_OECD_HIC %>%
  synthetic_control(
    outcome = `EN.ATM.CO2E.PC`, 
    unit = country,
    time = year,
    i_unit = "United Kingdom",
    i_time = 2001,
    generate_placebos = TRUE) %>%
  
  generate_predictor(  # Predictors (pre-treatment averages)
    time_window = 1980:2001,
    mean.preT = mean(mean.preT, na.rm = TRUE)) %>%
  
  generate_weights(
    optimization_window = 1980:2001,
    margin_ipop = 0.02,
    sigf_ipop = 7,
    bound_ipop = 6) %>%
  generate_control() # Generate synthetic control



# data_OECD_HIC_UMC
scm_data_OECD_HIC_UMC <- data_OECD_HIC %>%
  synthetic_control(
    outcome = `EN.ATM.CO2E.PC`, 
    unit = country,
    time = year,
    i_unit = "United Kingdom",
    i_time = 2001,
    generate_placebos = TRUE) %>%
  
  generate_predictor(  # Predictors (pre-treatment averages)
    time_window = 1980:2001,
    mean.preT = mean(mean.preT, na.rm = TRUE)) %>%
  
  generate_weights(
    optimization_window = 1980:2001,
    margin_ipop = 0.02,
    sigf_ipop = 7,
    bound_ipop = 6) %>%
  generate_control() # Generate synthetic control


```

## Plot Trends #1

```{r}
# data_OECD
p1_data_OECD <- scm_data_OECD %>%
  plot_trends() +
  labs(title = "UK vs Synthetic UK",
       y = "CO2 per capita (power sector)",
       x = "Year")

p1_data_OECD

# data_OECD_HIC
p1_data_OECD_HIC <- scm_data_OECD_HIC %>%
  plot_trends() +
  labs(title = "UK vs Synthetic UK",
       y = "CO2 per capita (power sector)",
       x = "Year")

p1_data_OECD_HIC

# data_OECD_HIC_UMC
p1_data_OECD_HIC_UMC <- scm_data_OECD_HIC_UMC %>%
  plot_trends() +
  labs(title = "UK vs Synthetic UK",
       y = "CO2 per capita (power sector)",
       x = "Year")

p1_data_OECD_HIC_UMC 
```

## TRY #2 

```{r}
# data_OECD
scm_data_OECD <- data_OECD %>%
  synthetic_control(
    outcome = EN.ATM.CO2E.PC,
    unit = country,
    time = year,
    i_unit = "United Kingdom",
    i_time = 2001,
    generate_placebos = TRUE) %>%
  generate_predictor(time_window = 1980, EN.ATM.CO2E.PC_1980 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1981, EN.ATM.CO2E.PC_1981 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1982, EN.ATM.CO2E.PC_1982 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1983, EN.ATM.CO2E.PC_1983 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1984, EN.ATM.CO2E.PC_1984 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1985, EN.ATM.CO2E.PC_1985 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1986, EN.ATM.CO2E.PC_1986 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1987, EN.ATM.CO2E.PC_1987 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1988, EN.ATM.CO2E.PC_1988 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1989, EN.ATM.CO2E.PC_1989 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1990, EN.ATM.CO2E.PC_1990 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1991, EN.ATM.CO2E.PC_1991 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1992, EN.ATM.CO2E.PC_1992 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1993, EN.ATM.CO2E.PC_1993 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1994, EN.ATM.CO2E.PC_1994 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1995, EN.ATM.CO2E.PC_1995 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1996, EN.ATM.CO2E.PC_1996 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1997, EN.ATM.CO2E.PC_1997 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1998, EN.ATM.CO2E.PC_1998 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1999, EN.ATM.CO2E.PC_1999 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 2000, EN.ATM.CO2E.PC_2000 = mean(EN.ATM.CO2E.PC)) %>%
  
  generate_weights(optimization_window = 1980:2001) %>%
  generate_control()

# data_OECD_HIC
scm_data_OECD_HIC <- data_OECD_HIC %>%
  synthetic_control(
    outcome = EN.ATM.CO2E.PC,
    unit = country,
    time = year,
    i_unit = "United Kingdom",
    i_time = 2001,
    generate_placebos = TRUE) %>%
  generate_predictor(time_window = 1980, EN.ATM.CO2E.PC_1980 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1981, EN.ATM.CO2E.PC_1981 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1982, EN.ATM.CO2E.PC_1982 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1983, EN.ATM.CO2E.PC_1983 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1984, EN.ATM.CO2E.PC_1984 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1985, EN.ATM.CO2E.PC_1985 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1986, EN.ATM.CO2E.PC_1986 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1987, EN.ATM.CO2E.PC_1987 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1988, EN.ATM.CO2E.PC_1988 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1989, EN.ATM.CO2E.PC_1989 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1990, EN.ATM.CO2E.PC_1990 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1991, EN.ATM.CO2E.PC_1991 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1992, EN.ATM.CO2E.PC_1992 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1993, EN.ATM.CO2E.PC_1993 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1994, EN.ATM.CO2E.PC_1994 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1995, EN.ATM.CO2E.PC_1995 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1996, EN.ATM.CO2E.PC_1996 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1997, EN.ATM.CO2E.PC_1997 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1998, EN.ATM.CO2E.PC_1998 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1999, EN.ATM.CO2E.PC_1999 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 2000, EN.ATM.CO2E.PC_2000 = mean(EN.ATM.CO2E.PC)) %>%
  
  generate_weights(optimization_window = 1980:2001) %>%
  generate_control()

# data_OECD_HIC_UMC
scm_data_OECD_HIC_UMC <- data_OECD_HIC_UMC %>%
  synthetic_control(
    outcome = EN.ATM.CO2E.PC,
    unit = country,
    time = year,
    i_unit = "United Kingdom",
    i_time = 2001,
    generate_placebos = TRUE) %>%
  generate_predictor(time_window = 1980, EN.ATM.CO2E.PC_1980 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1981, EN.ATM.CO2E.PC_1981 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1982, EN.ATM.CO2E.PC_1982 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1983, EN.ATM.CO2E.PC_1983 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1984, EN.ATM.CO2E.PC_1984 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1985, EN.ATM.CO2E.PC_1985 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1986, EN.ATM.CO2E.PC_1986 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1987, EN.ATM.CO2E.PC_1987 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1988, EN.ATM.CO2E.PC_1988 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1989, EN.ATM.CO2E.PC_1989 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1990, EN.ATM.CO2E.PC_1990 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1991, EN.ATM.CO2E.PC_1991 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1992, EN.ATM.CO2E.PC_1992 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1993, EN.ATM.CO2E.PC_1993 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1994, EN.ATM.CO2E.PC_1994 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1995, EN.ATM.CO2E.PC_1995 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1996, EN.ATM.CO2E.PC_1996 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1997, EN.ATM.CO2E.PC_1997 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1998, EN.ATM.CO2E.PC_1998 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 1999, EN.ATM.CO2E.PC_1999 = mean(EN.ATM.CO2E.PC)) %>%
  generate_predictor(time_window = 2000, EN.ATM.CO2E.PC_2000 = mean(EN.ATM.CO2E.PC)) %>%
  
  generate_weights(optimization_window = 1980:2001) %>%
  generate_control()

```


```{r}
scm_data_OECD %>%
  grab_unit_weights() %>%
  arrange(desc(weight))

scm_data_OECD_HIC %>% 
    grab_unit_weights() %>%
  arrange(desc(weight))

scm_data_OECD_HIC_UMC %>% 
    grab_unit_weights() %>%
  arrange(desc(weight)) %>% 
filter(weight > 0.01)
```
```{r}
scm_data_OECD %>% 
    plot_weights()

scm_data_OECD_HIC %>% 
    plot_weights()
    
scm_data_OECD_HIC_UMC %>% 
    plot_weights()
```

## Plot Trends #2 

```{r}
# data_OECD
p1_data_OECD <- scm_data_OECD %>%
  plot_trends() +
  labs(title = "UK vs Synthetic UK",
       y = "CO2 per capita (power sector)",
       x = "Year")

p1_data_OECD

# data_OECD_HIC
p1_data_OECD_HIC <- scm_data_OECD_HIC %>%
  plot_trends() +
  labs(title = "UK vs Synthetic UK",
       y = "CO2 per capita (power sector)",
       x = "Year")

p1_data_OECD_HIC

# data_OECD_HIC_UMC
p1_data_OECD_HIC_UMC <- scm_data_OECD_HIC_UMC %>%
  plot_trends() +
  labs(title = "UK vs Synthetic UK",
       y = "CO2 per capita (power sector)",
       x = "Year")

p1_data_OECD_HIC_UMC 
```

## Plot Gaps (Treatment Effect)
```{r}
scm_data_OECD_HIC_UMC %>%
  filter(.id == "United Kingdom", .placebo == 0) %>%
  pull(.synthetic_control) %>%
  .[[1]] %>% 
  mutate(gap = real_y - synth_y) %>%
  filter(time_unit >= 2001)
```


```{r}
# data_OECD
p2_data_OECD <- scm_data_OECD %>%
  plot_differences() +
  labs(title = "Gap: Observed - Synthetic")

p2_data_OECD

# data_OECD_HIC
p2_data_OECD_HIC <- scm_data_OECD_HIC %>%
  plot_differences() +
  labs(title = "Gap: Observed - Synthetic")

p2_data_OECD_HIC

# data_OECD_HIC_UMC
p2_data_OECD_HIC_UMC <- scm_data_OECD_HIC_UMC %>%
  plot_differences() +
  labs(title = "Gap: Observed - Synthetic")

p2_data_OECD_HIC_UMC 

```

## Placebo Tests

```{r}
# data_OECD
p3_data_OECD <- scm_data_OECD %>%
  plot_placebos() +
  labs(title = "Placebo Tests")

p3_data_OECD

# data_OECD_HIC
p3_data_OECD_HIC <- scm_data_OECD_HIC %>%
  plot_placebos() +
  labs(title = "Placebo Tests")

p3_data_OECD_HIC

# data_OECD_HIC_UMC
p3_data_OECD_HIC_UMC <- scm_data_OECD_HIC_UMC %>%
  plot_placebos() +
  labs(title = "Placebo Tests")

p3_data_OECD_HIC_UMC

```

## MSPE Ratio Plot
```{r}
scm_data_OECD %>%
  grab_significance() %>%  
  filter(unit_name == "United Kingdom")

scm_data_OECD_HIC %>%
  grab_significance() %>%  
  filter(unit_name == "United Kingdom")

scm_data_OECD_HIC_UMC %>%
  grab_significance() 

  filter(unit_name == "United Kingdom")
```
```{r}
scm_data_OECD_HIC_UMC %>%
  grab_significance() %>%
  filter(pre_mspe < 5 * min(pre_mspe[type == "Treated"])) %>%
  arrange(desc(mspe_ratio))
```



```{r}
# data_OECD
p4_data_OECD <- scm_data_OECD %>%
  plot_mspe_ratio() +
  labs(title = "MSPE Ratio (Permutation Inference)")

p4_data_OECD 

# data_OECD_HIC
p4_data_OECD_HIC <- scm_data_OECD_HIC %>%
  plot_mspe_ratio() +
  labs(title = "MSPE Ratio (Permutation Inference)")

p4_data_OECD_HIC

# data_OECD_HIC_UMC
p4_data_OECD_HIC_UMC <- scm_data_OECD_HIC_UMC %>%
  plot_mspe_ratio() +
  labs(title = "MSPE Ratio (Permutation Inference)")

p4_data_OECD_HIC_UMC 

```

## Balance Table

```{r}
# data_OECD
balance_data_OECD <- scm_data_OECD %>%
  grab_balance_table()

balance_data_OECD

# data_OECD_HIC
balance_data_OECD_HIC <- scm_data_OECD_HIC %>%
  grab_balance_table()

balance_data_OECD_HIC

# data_OECD_HIC_UMC
balance_data_OECD_HIC_UMC <- scm_data_OECD_HIC_UMC %>%
  grab_balance_table()

balance_data_OECD_HIC_UMC
```

# Finding difference in means with pre-treatment values- this is our main result
```{r}
scm_data_OECD_HIC_UMC %>%
    # Extract values from balance table
    grab_balance_table() %>%
    # Get differences between actual UK and synthetic and compare to donor differences
    mutate(
        synth_diff = `United Kingdom` - `synthetic_United Kingdom`,
        donor_diff = `United Kingdom` - donor_sample,
        year = as.integer(str_extract(variable, "\\d{4}"))
    ) %>%
    # Make into a tidy data format
    pivot_longer(
        cols = c(synth_diff, donor_diff),
        names_to = "type",
        values_to = "diff"
    ) %>%
    # Recode values to match those in paper result
    mutate(type = recode(type, "synth_diff" = "Synthetic control", "donor_diff" = "Unweighted sample")) %>%
    # Plot
    ggplot(aes(x = diff, y = factor(year), color = type)) +
    geom_point(size = 3) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    # Match colors used by paper figure
    scale_color_manual(values = c(
        "Synthetic control" = "#E07B6A",
        "Unweighted sample" = "#5BBCD6"
    )) +
    labs(
        x = "Difference in means with pre-treatment values in the UK",
        y = "Year",
        # Take out legend title the lazy way
        color = "",
        title = "Balance in weighted synthetic control and in unweighted sample"
    ) +
    # Get all the lines out
    theme_classic()
```

